---
title: "COVID19 Vaccines in Israel"
author: "Adi Sarid"
date: "2/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(glue)
library(gt)
library(gtsummary)
```

How is Israel doing in terms of population vaccinations?

Where are vaccinations adopted and where they are not going so well?

## Read the data 

```{r read the vaccinations data, message=FALSE, warning=FALSE}
vacci_raw <- read_csv("data/2021-02-03 - vaccinated_city_table_ver_0013.csv")
glimpse(vacci_raw)
```

Observations about how the data is arranged:

   * Wide format where the age group is noted in the header. 
   * The values are character (instead of numeric) since one possible value is "<15" (for privacy reasons statistical cells less than 15 are not reported).
   * Also, different headers for first dose and second dose.
   * According to the [documentation](https://data.gov.il/dataset/covid-19/resource/c70cff54-4e4c-47dc-b47c-7500f366319b?inner_span=True) only towns with pop. size > 2000 are reported.

Reshaping the file a bit into a long format, and getting rid of the characters. I will replace "<15" with 0.

```{r arranging the file}

cbs_data <- readxl::read_excel("data/bycode2018.xlsx") %>% 
   select(city_symbol, population)

vacci <- vacci_raw %>% 
  pivot_longer(cols = contains("_dose_"),
               names_to = "vaccine_type", 
               values_to = "vaccinated") %>% 
  mutate(vaccinated = if_else(vaccinated == "<15", "0.0", vaccinated)) %>% 
  mutate(vaccinated = parse_number(vaccinated)) %>% 
  mutate(dose = case_when(str_detect(vaccine_type, "first") ~ 1,
                          str_detect(vaccine_type, "second") ~ 2,
                          TRUE ~ -99)) %>% 
   mutate(age_group = str_remove(vaccine_type, "first_dose_|second_dose_")) %>% 
   left_join(cbs_data, by = c("CityCode" = "city_symbol"))

vacci_agg <- vacci %>% 
   filter(!is.na(population)) %>% 
   group_by(CityName, dose, Date, population) %>% 
   summarize(total_vaccinated = sum(vaccinated)) %>% 
   mutate(dose = paste0("dose", dose)) %>% 
   pivot_wider(names_from = dose, values_from = total_vaccinated) %>% 
   mutate(dose1_prcnt = dose1/population,
          dose2_prcnt = dose2/population)
      
vacci_agg %>% 
   filter(CityName %in% c("בני ברק",
                          "נתניה",
                          "תל אביב - יפו",
                          "חיפה",
                          "ירושלים")) %>% 
   ggplot(aes(x = Date, y = dose1_prcnt, color = CityName)) + 
   geom_line(size = 1) + 
   saridr::theme_sarid() + 
   ylab("Population [%]") + 
   ggtitle("\u202bחיסון מנה ראשונה של COVID19 בישראל - רשויות נבחרות") + 
   scale_y_continuous(labels = scales::percent_format(1)) + 
   guides(color = guide_legend("עיר")) +
   labs(caption = "\u202bמקור הנתונים: מאגר COVID-19 של משרד הבריאות מתחסנים על פי יישוב\n\u202bויז'ואליזציה מכון שריד: https://www.sarid-ins.co.il") + 
   theme(plot.title = element_text(hjust = 1))

```